\documentclass[a4paper, 11pt]{report}

\input{packages}
\input{commands_and_environments}

%\includeonly{p3/p3}

\bibliography{bib}
\addbibresource{bib.bib}

\begin{document}

\chapter{TODO-Lists}
\label{chap:todo}

\section{Current questions}
\label{chap:current-questions}

\begin{itemize}
\item P3: Is it possible, that there are -- at one level of the optimal snapshot DAG -- multiple snapshots containing the same intree, but a different set of scheduled tasks? \remark{Probably not w.l.o.g.}
\item For each snapshot resulting from an optimal schedule, at least one top-most task is scheduled? \remark{Seems so.}
\item If for an intree only non-top tasks are scheduled, you can exchange one of the non-top tasks with a top-tasks and obtain a better run time?\remark{Seems so}
\item ``Topmost singles:'' Are topmost tasks, whose successors have \emph{exactly one} predecessor, always scheduled?\done{No! See chapter ``suboptimal strategies'' (number of topmost tasks).}
\item Are there intrees where we have to schedule topmost tasks of \emph{three different (root-)subtrees? (Root-subtree: Subtree rooted at a direct predecessor of the root.)}\done{Yes. Siehe suboptimal-Kapitel, irgendein non-HLF-optimal-Beispiel.}
\item If a tree $T$ is non-HLF-optimal, can there a be a supertree $S$ of this tree ($T\subseteq S$) such that $S$ is HLF-optimal? \done{Yes, refer to section \ref{sec:properties-optimal-schedules-no-implications}.}
\item Are chain collections optimally scheduled by HLF?\done{Possibly so. Check proof once again.}
\item Let's assume we force the \emph{first step} to be HLF. Is then the best possible solution with exactly this HLF-beginning again HLF?\done{No -- consider the non-HLF examples!}
\end{itemize}

\section{Other TODOs}

\begin{itemize}
\item Program does incorrect output of initially scheduled tasks.
\item TikZ export very slow due to \texttt{generate\_tikz\_nodes}.
\item Appendix containing all tree sequences for which optimal schedule is strictly non-HLF.
\item Improve program by using arrays instead of maps (only for canonical snapshots). \done{Works for canonical snapshots, ``normal ones'' broken.} \todo{Correct it!}
\end{itemize}

%\input{../0011123669hlf.tex}
%\input{examples/p2/001112566_with_condensed.tex}
%\input{../maxtrees}
%\input{../default}
%\include{../default_optimized}
%\include{../default_profile}
%\include{../default_chain}

% here, the document actually starts to get useful

%\part{Introduction}
\include{theory/theory}

%\part{Warmup: The two-processor case}
\include{p2/p2}

%\part{The three processor case}
\include{p3/suboptimal_strategies}
\include{p3/p3}

%\part{Implementation details}
\include{additional_algs/additional_algs}
\include{benchmark/benchmark}

\include{first/first}

\appendix{}

%\include{counterexamples/counterexamples}

%\bibliographystyle{alphabetic}
\nocite{*}
\printbibliography[notkeyword=oeis,title={Bibliography}]
%\printbibliography[title={Number sequences},keyword=oeis]

\end{document}
